---
title: "lab_1"
authror: "Akiva Finkelstein & Amit Yaron"
output:
  html_document: default
  pdf_document: default
  word_document: default

---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE, message=FALSE, include=FALSE, echo=FALSE}
library(nycflights13)
library(dplyr)
library(maps)
library(usmap)
library(ggplot2)
library(plotly)
library(gridExtra )
library(GGally)
```
#Q1
#Graph(I)
1) The first plot wants to answer the question of the percentage of flights,\n departing from Denver airport, delayed 15 minutes and more.\n
2)To our opinion the 
the plot shows the information well. The colors of each delay category shows to\n
which destination the proportion of delayed flights was high/low. Which can be\n very useful for optimizing scheduling.\n 
3) The main question that arises is what could be the reason for all these\n 
delays. It would be useful to check the connection to the airline carrier or\n
to see if there are spesific times that have more delays.\n
4)We believe that the graph would be easier to understand if the lines on the\n graph\n would thinner. It wouldn't have blocked such a large area of the map,\n blocking the many state abbreviations.


#Graph (II)
1) 
The graph shows the cycles of the number of flights per day, and the number of\n 
delayed flights from all flights every day.The graph answer the question among \n
all the departing flight how many of the departing flights are with a delay of \n
at least 15 minutes.

2) Yes we can see clearly each day how many out of the total flights  had a delay.\n
since both total flights and delayed flights are potted together it gives us a\n
good picture of how many flights are delayed with respect to all flights in general. 
3)Yes. IF we take a look at the graph from 31 August  to 2 November, we can see\n
that departing flights are steady at around 700 per day.On the other hand we can\n
see the graph of the delayed flights does fluctuate quite. Perhaps the amount of\n
departing flights each day does not  effect the amount of delayed flights.\n
4)We believe the information would be better understood if the plot would be split\n
in to two separate plots. it is simply easier on the eyes. 

#Q2
#Graph of USA map with with connecting routs showing the percetage of departure \ndelay from JFK airport.  
```{r,  warning=FALSE, message=FALSE, include=FALSE, echo=FALSE}
flights_data <- flights
flights_data <- flights_data %>% filter(origin == "JFK")

flights_data$long_delay <- ifelse(flights_data$dep_delay >= 15,1,0) #adding long delay column

flights_data <- flights_data %>% group_by(origin,dest,)%>%summarize(num_flights = n(),
                                                                num_delays = sum(long_delay,na.rm = TRUE))
#percent of delays 
flights_data$delay_percent <-as.numeric(flights_data$num_delays / flights_data$num_flights)
#breaking into category 
flights_data$delay_deg <- cut(flights_data$delay_percent, breaks =c(-0.5, 0.1,0.15,0.2,0.25,Inf),labels = c("<=10%","10%-15%", "15%-20%", "20%-25%", ">25%"))

flights_data[,c("delay_deg")][is.na(flights_data[,c("delay_deg")])] <- 0


# getting lat and long of airports destination
a <- airports
a <- a %>% rename(dest = faa) 
a1 <- data.frame(a$dest, a$lat, a$lon)
a1 <- a1 %>% rename(dest =a.dest) %>% rename(dest_lon = a.lon) %>%
  rename(dest_lat = a.lat)

flights_data <- inner_join(flights_data, a1, by = "dest") 
a1 <- a1 %>% rename(origin = dest)
# getting lat and long of airports destination departures 
flights_data <- inner_join(flights_data, a1, by = "origin")

flights_data<- flights_data %>% rename(dest_lat = dest_lat.x) %>%
  rename(dest_lon = dest_lon.x) %>% rename(dep_lat = dest_lat.y) %>% 
  rename(dep_lon = dest_lon.y)




lable_state <- data.frame(state.abb,state.center)#state coordinates
lable_state <- lable_state %>% filter(state.abb != "AK") %>% filter(state.abb != "HI") 
usa <- map_data("state") #map data


```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
#map


gg4 <- ggplot()+
  geom_polygon( data=usa, aes(x=long, y=lat, group=group),
                color="gray", fill="white" )+
  geom_text(data = lable_state, aes(x, y, label = state.abb), size = 2)+
  geom_segment(data = flights_data,aes(x = dep_lon, y = dep_lat, xend = dest_lon,
                                       yend = dest_lat, color = delay_deg), size =rel(0.2))+
  coord_map(xlim = c(-127,-55), ylim=c(22,50))+
  geom_point(data=flights_data, aes(dest_lon, dest_lat),
             color = "gray",size = rel(1), shape=1)+
  labs(title = "% of Flights Delayed > 15 min", subtitle = "Airport = JFK       Year = 2013")+
theme(axis.title.y = element_blank(),
      axis.title.x=element_blank(), 
      plot.background = element_rect(color = "black"),
      panel.background=element_rect(fill = "white"),
      axis.ticks = element_blank(),
      axis.text=element_blank(),
      panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      legend.key = element_rect(fill = "white"),
      legend.title = element_blank(),
      legend.position = c(0.07,0.2),
      plot.margin =  margin(1,0,0,0,"cm"),
      plot.title = element_text(size = rel(1.5), hjust = 0.5,vjust = 2, face = "bold"),
      plot.subtitle = element_text(size = rel(1), hjust = 0.5, color="grey 39", face = "bold"))
  
gg4
 
```

 
b)
#graphic summarizing the flight volume and fights delayed,\n by date and showing weekly cycles.
```{r,message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
#Manipulating data tables


flights_data <- flights #Upload the data 
flights_delay <-filter(flights_data,dep_delay >= 15) #table with 15> delay 
#convert (year month day) -> date
flights_data$date <- as.Date(with(flights_data, paste(year, month, day,sep="-")), "%Y-%m-%d")
flights_delay$date <- as.Date(with(flights_delay, paste(year, month, day,sep="-")), "%Y-%m-%d")


#count the number of flight and number of flight with delay>=15
t<-flights_data %>% group_by(date,year) %>% summarise(num_flights = n())
h<-flights_delay %>% group_by(date,year) %>% summarise(flights_delay = n())



#merge the data to one data frame
gdata<-merge(t,h)
#select points
b1<-gdata[gdata$date=="2013-07-04",]
b2<-gdata[gdata$date=="2013-11-28",]
b<-rbind(b1,b2)

```



```{r ,fig2, fig.height = 4, fig.width = 7, fig.align = "center", echo=FALSE}
#plot the graph
#Graph with All Flight
fig <- plot_ly(
  type = "scatter",
  x = as.Date(gdata$date, format= "%Y-%m-%d"), 
  y = gdata$num_flights,
  name = 'All Flight(scheduled for departure) ',
  mode = "lines+markers",
)


#The min Points in the All Flight Graph
fig<- fig %>% add_trace(y= ~b$num_flights,x=~b$date, 
                        name = 'FEWER FLIGHTS ON HLIDAYS'
                        ,type='scatter'
                        ,mode='markers'
                        ,size=10)
fig <- fig %>% layout(xaxis = list(title = ""))



#Graph with Late Flights(departure delayed>15 min)
fig <- fig %>%
  layout(
    title = "WEEKLY CYCLES-Temporal Effects",
    xaxis = list(
      type = "date",
      range=c('2013-01-01', '2014-01-01,date_breaks = "1 month"')
    )
  )
fig<- fig %>% add_trace(y= ~gdata$flights_delay, name = 'Late Flights(departure delayed>15 min)',type = 'scatter', mode = 'lines+markers')
fig <- fig %>% layout(yaxis = list (title = "Flight Per Days"))
fig <- fig %>% layout(legend = list(orientation = 'h'))




fig

```


#Q3
```{r, warning=FALSE, message=FALSE, include=FALSE, echo=FALSE}

# percent of flights delayed by carrier
flights_data3 <- flights

flights_data3$long_delay <- ifelse(flights_data3$dep_delay > 15,1,0) #adding long delay column

flights_data3 <- flights_data3 %>% group_by(carrier)%>%summarize(num_flights = n(),
                                                            num_delays =sum(long_delay,na.rm=TRUE))

flights_data3$percent_delay <- round((flights_data3$num_delays / flights_data3$num_flights)*100,1)
flights_data3$num_flights<- round(flights_data3$num_flights/1000,2)
```


```{r, warning=FALSE, message=FALSE, echo=FALSE}
g3 <- ggplot(flights_data3, aes(carrier, percent_delay)) + geom_bar(stat = "identity",fill="purple") +
scale_x_discrete("Airline Carrier") +
  scale_y_continuous(expand = c(0.01,0.01))+
  labs(title = "% of  Arrival Delays by Carrier", subtitle = "percetage of flights 
  delayed 15 Min or more",y = "precent of arrival\ndelayed flights") +
  geom_text(aes(label=percent_delay), vjust=1.6, size=3.0, color = "red")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = rel(1.2)),axis.text.x = element_text(angle = 90),
        plot.subtitle = element_text(hjust = 0.5, size = rel(0.8)),
        axis.title.y = element_text(angle = 90, size = rel(1),),
        axis.title.x = element_text(angle = 0,size = rel(1)),)
  
g3.1 <- ggplot(flights_data3, aes(carrier, num_flights)) + geom_bar(stat = "identity",fill="steelblue") +
scale_x_discrete("Airline Carrier") +
  scale_y_continuous(expand = c(0.01,0.01))+
  labs(title = "Number of Flights by Carrier",subtitle = "Number of flights in thousands ", y = "Number of Flights") +
  geom_text(aes(label=num_flights), vjust=1.5, size=2.8, color = "red")+
  theme_bw()+
  theme(plot.title = element_text(hjust = 0.5, size = rel(1.2)),axis.text.x = element_text(angle = 90),
        plot.subtitle = element_text(hjust = 0.5, size = rel(0.8)),
        axis.title.y = element_text(angle = 90, size = rel(1)),
        axis.title.x = element_text(angle = 0,size = rel(1)),)



grid.arrange(g3.1, g3, 
             ncol = 2, nrow = 1)
```

We can see from both plots how many flights in total each had and the percentage\n
of flights with arrival delays of more then 15 minutes. Plotting these graphs\n together gives us a good picture on how well each airline does with minimizing delayed flights. For \n
instance we can see that UA has the most flights in total but does not have the\n
highest percentage of flights being delayed. Hence we can say that that amount \n
of flights are not necessarily connected to high percentage of delays.  

```{r,message=FALSE, warning=FALSE, include=FALSE, echo=FALSE}
#Manipulating data tables
weather_data<-weather
weather_data$date<-as.Date(with(weather_data, paste(year, month, day,sep="-")), "%Y-%m-%d")#put date in the weather_date
planes_data<-planes
planes_seat100<-filter(planes_data,seats <= 100) #Choose only small planets <=100 seat 

Q <- flights_data%>% filter(flights_data$tailnum == planes_seat100$tailnum)#Choose only flights data of small seat <=100 it include date ! 

adata <- merge(Q,weather_data,by = c("origin","year","month","day","date","hour","time_hour"))



```


```{r,message=FALSE, warning=FALSE, echo=FALSE}
#plot the graph


analysis<-data.frame(adata$dep_delay,adata$arr_delay,adata$temp,adata$wind_speed,adata$humid,adata$visib)
names(analysis)[1]<-"dep_delay"
names(analysis)[2]<-"arr_delay"
names(analysis)[3]<-"temp"
names(analysis)[4]<-"wind_speed"
names(analysis)[5]<-"humid"
names(analysis)[6]<-"visib"
#view(analysis)


g9<-ggscatmat(analysis,alpha=0.8) +
   ggtitle("Joint distribution of continuous variables scatterplot matrix") + 
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=0, size=4), axis.text.y = 
          element_text(angle=0, vjust=1, hjust=0, size=5))+
  scale_x_continuous(expand = c(0.1,0.1))
g9
```

In this graph we Try to understand which variables  related to weather effect\n
the departure delay and how they effect each other.
We focused on aircraft with up to 100 seats.
we choose the weather variable : humid,temperature,wind speed , viability (how far we can see on miles).
The Statistical  measurement we work with are join distribution correlation matrix and  approximate the histogram of the data.
 
we can see we have strong  a strong connection between  arrivals delay and departure delay
the correlation there is 0.94 also the join distribution graph looks like  line\n
and not like random cloud.

Also we can see some little  connection between humid and arrivals delay the correlation there is 0.24 
and in the  join distribution graph we can that most of the point found mostly in the left area of the graph.

In addition There is An interesting connection between visib and humid the\n
correlation there is -0.54. Also the  joint distribution graph looks like A straight line.

Finally the wind_speed is distrusted like Normal distribution
And departure delay,arrivals delay came from the same distribution family with different parameters.


#Q4
We will check the null hypothesis:The proportion of delayed flights per month\n
is independent across months.\n
First we plot the real data in a bar pot. Then we will simulate plots from the\n
real data, and try to identify the real data in the lineup. WE will plot 19 simulated plots to\n
get a significance level of 95%. If we can detect the true data then we will reject\n
the null hypothesis   
```{r, warning=FALSE, message=FALSE, include=FALSE, echo=FALSE}
flights_data4 <- flights
flights_data4 <- flights_data4 %>% filter(origin == "JFK")

flights_data4$long_delay <- ifelse(flights_data4$dep_delay > 15,1,0) #adding long delay column

flights_data4 <- flights_data4 %>% group_by(month)%>%summarize(num_flights = n(),
                                                        num_delays = sum(long_delay,na.rm = TRUE))
flights_data4$prop_delay <- flights_data4$num_delays / flights_data4$num_flights

flights_data4$month <- month.abb # converting to month.abb format
```

```{r,warning=FALSE,message=FALSE, echo=FALSE}
                                                                                    
# bar plot
g1<-  ggplot(data = flights_data4) +
  geom_col(aes(x = month, y = prop_delay)) +
  scale_x_discrete(limits = month.abb) +
  scale_y_continuous(expand = c(0.01,0.01), breaks = seq(0,0.5,0.025))+
   labs(title = "True plot of the null hypothesis",
        subtitle = "Null Hypothesis: The proportion of delayed flights\n
        per month is independent across months",
        y = "precent of\n delayed flights\n per month") +
   
   theme_bw()+
   theme(plot.title = element_text(hjust = 0.5, size = rel(1.5)),
         plot.subtitle = element_text(hjust = 0.5, size = rel(0.8)),
         axis.title.y = element_text(angle = 0, size = rel(0.75),),
         axis.title.x = element_text(size = rel(0.75)),)
g1 
```



```{r, warning=FALSE, message=FALSE, echo=FALSE}

x_rand <- as.data.frame(replicate(19, sample(flights_data4$prop_delay)))
x_rand$month <- month.abb
x_rand <- as.data.frame(x_rand)

par(mfrow=c(4,5),oma = c(2, 2, 2, 2))
par(mar=rep(2,4))
barplot(height = flights_data4$prop_delay, names.arg =x_rand$month, main = "1",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V1, main = "2",names.arg =x_rand$month ,border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V2, names.arg = x_rand$month, main = "3",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V3, names.arg = x_rand$month, main = "4",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V4, names.arg = x_rand$month, main = "5",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V5, names.arg = x_rand$month, main = "6",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V6, names.arg = x_rand$month, main = "7",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V7, names.arg = x_rand$month, main = "8",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V8, names.arg = x_rand$month, main = "9",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V9, names.arg = x_rand$month, main = "10",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V10, names.arg = x_rand$month, main = "11",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V11, names.arg = x_rand$month, main = "12",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V12, names.arg = x_rand$month, main = "13",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V13, names.arg = x_rand$month, main = "14",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V14, names.arg = x_rand$month, main = "15",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V15, names.arg = x_rand$month, main = "16",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V16, names.arg = x_rand$month, main = "17",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V17, names.arg = x_rand$month, main = "18",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V18, names.arg = x_rand$month, main = "19",border="#69b3a2",las=2,cex.names=0.5)
barplot(height = x_rand$V19, names.arg = x_rand$month, main = "20",border="#69b3a2",las=2,cex.names=0.5)
mtext("Eye training: Delay Flight Proportion ", outer = TRUE, cex = 1)
mtext("Month",at=-0.07, adj=0, cex=0.7,outer=TRUE,side = 3, line = 0)



```


#Lineup
when examining the lineup it is  hard to identify the real data from the simulated\n
data, with out having prior knowledge of the true data. Thus we can not reject \n
the null hypothesis i.e we can't say that the proportion of delayed flights per\n
month is independent across months.
** The true data is in plot 1 **


